---
title: "Earthquakes"
author: "Alicia Brown"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8,
  fig.height=5
)
```

The `earthquakes` library allows users to-

- Pass raw earthquake data to `eq_clean_data` and be returned date, city and numeric coordinates
- Visualize earthquakes over time
- Visualize earthquakes geospatially
- View annotations

## eq_clean_data
1. Constructs date from `YEAR`, `MONTH`, `DAY` (defaults NA to 1 for MONTH and DAY and excludes BCE events)
2. Sets Latitude and Longitude as numbers (doubles) for use by mapping libraries
3. Extracts city name from full `LOCATION_NAME` value and format as Full Title Case
```{r summaryData}
library(earthquakes)
library(dplyr)
data <- eq_clean_data(raw)
summary(data[c('DATE','LONGITUDE','LATITUDE','LOCATION_CITY')])

head(data[,c('DATE','LONGITUDE','LATITUDE','LOCATION_CITY')])
```

## eq_clean_location
1. `eq_clean_data` passes `COUNTRY` and `LOCATION_NAME` like _"ITALY: POMPEI"_
2. Country and colon are removed
3. Full Title Case applied
```{r cleanLocationCity}
example_country <- "ITALY"
example_location <- "ITALY: POMPEI"
clean_location <- eq_location_clean(example_country, example_location)
clean_location
```

## geom_timeline()
* visualize earthquakes over time
```{r geomtimeline, eval=T}
test_g <- geom_timeline(data, "2000-01-01", "2018-01-01", "NEW ZEALAND")
test_g
```
## geom_timeline()
* visualize earthquakes over time with a label for Location City of earthquake
```{r geomtimelinelable, eval=T}
test_g <- geom_timeline_label(data, "2000-01-01", "2018-01-01", "NEW ZEALAND",7)
test_g
```

## eq_map(data, annot_col)
* Visualize geographically earthquakes by magnitude with annotated label passed to function
```{r leafletViz}
test_m <- eq_clean_data(raw) %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000)
eq_map(test_m, "DATE")
```
