---
title: "Earthquakes"
author: "Alicia Brown"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `earthquakes` library allows users to-

- Pass raw earthquake data to `eq_clean_data` and be returned date, city and numeric coordinates
- Visualize earthquakes over time
- Visualize earthquakes geospatially
- View annotations

## eq_clean_data
1. Constructs date from `YEAR`, `MONTH`, `DAY` (defaults NA to 1 for MONTH and DAY and excludes BCE events)
2. Sets Latitude and Longitude as numbers (doubles) for use by mapping libraries
3. Extracts city name from full `LOCATION_NAME` value and format as Full Title Case
```{r summaryData}
library(earthquakes)

data <- eq_clean_data(raw)
summary(data[c('DATE','LONGITUDE','LATITUDE','LOCATION_CITY')])

head(data[,c('DATE','LONGITUDE','LATITUDE','LOCATION_CITY')])
```

## eq_clean_location
1. `eq_clean_data` passes `COUNTRY` and `LOCATION_NAME` like _"ITALY: POMPEI"_
2. Country and colon are removed
3. Full Title Case applied
```{r cleanLocationCity}
example_country <- "ITALY"
example_location <- "ITALY: POMPEI"
clean_location <- eq_location_clean(example_country, example_location)
clean_location
```
